Project 2

### 2018-09-30 Project2/report/Project2.tex
Jacobi's method complete. Starting quantum dots
Middle of one electron

\title{A spectacular title}

@@

\begin{abstract}
\end{abstract}
\keywords{cosmic microwave background --- cosmology: observations --- methods: statistical}

@@

\subsection{Buckling beam}
Considering a beam that is fastened in both ends which moves up and down when a force F is exerted on it can be described by a differential equation
%
\begin{gather}\label{eq:du2/dx2}
    R \frac{d^2 u(x)}{dx^2} = -Fu(x),
\end{gather}
%
where $u(x)$ is the displacement of the beam in the y-direction and $R$ is a constant that is defined by the properties of the beam, such as the rigidity. Since the beam, with length L, is fastened in both ends, the displacement at the end points $u(0) = u(L) = 0$, so we have Dirichlet boundary condition as in the previous paper \cite{1}, and the sum of outer forces exerted on the beam is zero. Introducing the dimensionless variable $\rho = x/L \Rightarrow x = \rho L$, where $\rho \in [0,1]$. Introducing this into Eq. \eqref{eq:du2/dx2} gives
%
\begin{gather*}
    \frac{R}{L}\frac{d^2 u(\rho)}{d\rho^2} = - FLu(\rho)
    \Rightarrow \frac{d^2 u(\rho)}{d\rho^2} = -\frac{FL^2}{R}u(\rho).
\end{gather*}
%
Substituting $\lambda = FL^2/R$ transforms the problem to an eigenvalue problem, with discretized $\lambda$'s as eigenvalues with corresponding eigenvectors. The methods presented in the previous paper regarding the solution of a differential equation will be applied to this problem. The left hand size of the equation to solve,
%
\begin{gather}\label{eq:du/dx=lambda}
    \frac{d^2 u(\rho)}{dx^2} = -\lambda u(\rho)
\end{gather}
%
can be approximated as
%
\begin{gather*}
    \frac{d^2 u(\rho)}{dx^2} = \frac{u(\rho + h) - 2u(\rho) + u(\rho - h)}{h^2} +  \mathcal{O}(h^2),
\end{gather*}
%
where the step size $h = \left(\rho_{max} - \rho_{min}\right)/n$, which is the total length of the beam divided by the resolution of the calculations n. The maximum and minimum value of the dimensionless length variable $\rho$ is at the end points of the beam, such that $\rho_{max} = 1$ and $\rho_{min} = 0$. Following the method from the previous paper the discretized equation to solve is
%
\begin{gather*}
    -\frac{u_{i+1} - 2u_i + u_{i-1}}{h^2} = \lambda u_i.
\end{gather*}
Which by the method of the previous paper can be transferred to a matrix eigenvalue problem;
%
\begin{equation*}
    \begin{bmatrix} d& a & 0   & 0    & \dots  &0     & 0 \\
                                a & d & a & 0    & \dots  &0     &0 \\
                                0   & a & d & a  &0       &\dots & 0\\
                                \dots  & \dots & \dots & \dots  &\dots      &\dots & \dots\\
                                0   & \dots & \dots & \dots  &a  &d & a\\
                                0   & \dots & \dots & \dots  &\dots       &a & d\end{bmatrix} 
                                 \begin{bmatrix} u_1 \\ u_2 \\ u_3 \\ \dots \\ u_{N-2} \\ u_{N-1}\end{bmatrix} = \lambda \begin{bmatrix} u_1 \\ u_2 \\ u_3 \\ \dots \\ u_{N-2} \\ u_{N-1}\end{bmatrix} . 
\label{eq:matrixse} 
\end{equation*}
%
Where a and b are defined as $a = -1/h^2$ and $b = 2/h^2$. For this eigenvalue problem the analytical eigenvalues are
%
\begin{gather}\label{eq:eigenvals_analytical}
    \lambda_j = d + 2a\cos\left(\frac{j\pi}{n+1}\right)
\end{gather}
for $j = 1,2,...,n-1$.

@@

@@

>>>>

\title{Project 2}

@@

\begin{abstract}
\end{abstract}
\keywords{buckling beam --- quantum dots --- harmonic oscillator- and Coulomb potentials --- methods: scaling of differential equations, Jacobi's method}

@@

\subsection{Buckling beam}
Considering a beam that is fastened in both ends which moves up and down when a force F is exerted on it can be described by a differential equation
%
\begin{gather}\label{eq:du2/dx2}
    R \frac{d^2 u(x)}{dx^2} = -Fu(x),
\end{gather}
%
where $u(x)$ is the displacement of the beam in the y-direction and $R$ is a constant that is defined by the properties of the beam, such as the rigidity. Since the beam, with length L, is fastened in both ends, the displacement at the end points $u(0) = u(L) = 0$, so we have Dirichlet boundary condition as in the previous paper \cite{1}, and the sum of outer forces exerted on the beam is zero. Introducing the dimensionless variable $\rho = x/L \Rightarrow x = \rho L$, where $\rho \in [0,1]$. Introducing this into Eq. \eqref{eq:du2/dx2} gives
%
\begin{gather*}
    \frac{R}{L}\frac{d^2 u(\rho)}{d\rho^2} = - FLu(\rho)
    \Rightarrow \frac{d^2 u(\rho)}{d\rho^2} = -\frac{FL^2}{R}u(\rho).
\end{gather*}
%
Substituting $\lambda = FL^2/R$ transforms the problem to an eigenvalue problem, with discretized $\lambda$'s as eigenvalues with corresponding eigenvectors. The methods presented in the previous paper regarding the solution of a differential equation will be applied to this problem. The left hand size of the equation to solve,
%
\begin{gather}\label{eq:du/dx=lambda}
    \frac{d^2 u(\rho)}{dx^2} = -\lambda u(\rho)
\end{gather}
%
can be approximated as
%
\begin{gather*}
    \frac{d^2 u(\rho)}{dx^2} = \frac{u(\rho + h) - 2u(\rho) + u(\rho - h)}{h^2} +  \mathcal{O}(h^2),
\end{gather*}
%
where the step size $h = \left(\rho_{max} - \rho_{min}\right)/n$, which is the total length of the beam divided by the resolution of the calculations n. The maximum and minimum value of the dimensionless length variable $\rho$ is at the end points of the beam, such that $\rho_{max} = 1$ and $\rho_{min} = 0$. Following the method from the previous paper the discretized equation to solve is
%
\begin{gather*}
    -\frac{u_{i+1} - 2u_i + u_{i-1}}{h^2} = \lambda u_i.
\end{gather*}
Which by the method of the previous paper can be transferred to a matrix eigenvalue problem;
%
\begin{equation*}
    \begin{bmatrix} d& a & 0   & 0    & \dots  &0     & 0 \\
                                a & d & a & 0    & \dots  &0     &0 \\
                                0   & a & d & a  &0       &\dots & 0\\
                                \dots  & \dots & \dots & \dots  &\dots      &\dots & \dots\\
                                0   & \dots & \dots & \dots  &a  &d & a\\
                                0   & \dots & \dots & \dots  &\dots       &a & d\end{bmatrix} 
                                 \begin{bmatrix} u_1 \\ u_2 \\ u_3 \\ \dots \\ u_{N-2} \\ u_{N-1}\end{bmatrix} = \lambda \begin{bmatrix} u_1 \\ u_2 \\ u_3 \\ \dots \\ u_{N-2} \\ u_{N-1}\end{bmatrix} . 
\label{eq:matrixse} 
\end{equation*}
%
Where a and b are defined as $a = -1/h^2$ and $b = 2/h^2$. For this specific eigenvalue problem the analytical eigenvalues are
%
\begin{gather}\label{eq:eigenvals_analytical}
    \lambda_j = d + 2a\cos\left(\frac{j\pi}{n+1}\right)
\end{gather}
for $j = 1,2,...,n-1$.

\subsection{Quantum dots, one electron}
The method derived in the section for the buckling beam is general for a problem that can be described by a differential equation on the form of \eqref{eq:du/dx=lambda}. The method can be extended to a quantum mechanical problem, if the problem can be transformed into a differential equation as mentioned. The radial part of the Schr$\ddot{\mathrm{o}}$dinger equation with the orbital angular momentum number $l = 0$ is
%
\begin{equation*}
    -\frac{\hbar^2}{2m}\left(\frac{1}{r^2}\frac{d}{dr}r^2\frac{d}{dr}\right)R(r) + V(r)R(r) = ER(r),
\end{equation*}
%
where $r \in [0,\infty)$, the potential $V(r)$ is the harmonic oscillator potential $V(r) = (1/2)kr^2$, where $k = mw^2$ and $\omega$ is the oscillator frequency. The allowed energies are 
%
\begin{equation*}
    E_{n0} = \left(2n + \frac{3}{2}\right),
\end{equation*}
%
for  n = 0, 1, 2,.... Substituting $R(r) = u(r)/r$:
%
\begin{equation*}
    -\frac{\hbar^2}{2m}\frac{1}{r^2}\frac{d}{dr}r^2 \frac{d}{dr}\frac{u(r)}{r} + V(r)\frac{u(r)}{r} = E\frac{u(r)}{r}.
\end{equation*}
%
Evaluating the derivatives in the above equation:
%
\begin{align*}
    \frac{1}{r^2}\frac{d}{dr}r^2\frac{d}{dr}\frac{u(r)}{r} &=  \frac{1}{r^2}\frac{d}{dr}r^2\left(\frac{\frac{du}{dr}r - u}{r^2}\right) \\
    &= \frac{1}{r^2}\frac{d}{dr}\left(\frac{du}{dr}r - u\right) \\
    &= \frac{1}{r^2}\left(\frac{d^2u}{dr^2}r + \frac{du}{dr} - \frac{du}{dr}\right) \\
    &= \frac{1}{r}\frac{d^2u}{dr^2},
\end{align*}
%
Inserting the above and multiplying by r yields
%
\begin{equation*}
    -\frac{\hbar^2}{2m}\frac{d^2u}{dr^2} + V(r)u(r) = Eu(r).
\end{equation*}
%
where the boundary conditions of $u$ are $u(0) = u(\infty) = 0$. Defining a dimensionless variable $\rho = r/\alpha$ transforms the potential to $V(\rho) = (1/2)k\alpha^2\rho^2$, and the Schr$\ddot{\mathrm{o}}$dinger's equation becomes
%
\begin{equation*}
    -\frac{\hbar}{2m\alpha^2}\frac{d^2u(\rho)}{d\rho^2} + \frac{1}{2}k\alpha^2\rho^2u(\rho) = Eu(\rho).
\end{equation*}
%
Multiplying the above equation by $2m\alpha^2 / \hbar^2$
%
\begin{equation*}
    -\frac{d^2u(\rho)}{d\rho^2} + \frac{mk}{\hbar^2}\alpha^2 \rho^2u(\rho) = \frac{2m\alpha^2}{\hbar^2}
\end{equation*}

@@

\subsection{Jacobi's method}

\subsection{Quantum dots, single electron}

\subsection{Quantum dots, two electrons}

@@

%We have managed to reduce a 3 dimentional two body problem to a analytically solvable problem by changing coordinate system from euclidian to radial, and analyizing CM-system(?)
